\begin{thebibliography}{9}
	\bibitem{root}
	Ahuja, R. K., O. Ergun, J. B. Orlin, A. P. Punnen. 2002. A survey of very large scale neighborhood search techniques. Discrete Appl. Math. 123 75–102.

	\bibitem{root}
	Bent, R., P. Van Hentenryck. 2003a. A two-stage hybrid algorithm for pickup and delivery vehicle routing problems with time windows. Lecture Notes in Computer Science, Vol. 2833. Proc. Internat. Conf. Constraint Programming (CP-2003), 123–137.

	\bibitem{root}
	Bent, R., P. Van Hentenryck. 2003b. A two-stage hybrid algorithm for pickup and delivery vehicle routing problem with time windows. Appendix available at http://www.cs.brown.edu/ people/rbent/pickup-appendix.ps.

	\bibitem{root}
	Bent, R., P. Van Hentenryck. 2004. A two-stage hybrid local search for the vehicle routing problem with time windows. Transporta- tion Sci. 38(4) 515–530.

	\bibitem{root}
	Cordeau, J.-F., G. Laporte, A. Mercier. 2001. A unified tabu search heuristic for vehicle routing problems with time windows. J. Oper. Res. Soc. 52 928–936.
	Cormen, T. H., C. E. Leiserson, R. L. Rivest, C. Stein. 2001. Intro- duction to Algorithms, 2nd ed. MIT Press, Boston, MA.

	\bibitem{root}
	Desaulniers, G., J. Desrosiers, A. Erdmann, M. M. Solomon, F. Soumis. 2002. The VRP with pickup and delivery. P. Toth, D. Vigo, eds. The Vehicle Routing Problem. SIAM Mono- graphs on Discrete Mathematics and Applications, Vol. 9. SIAM, Philadelphia, PA, 225–242.

	\bibitem{root}
	Dumas, Y., J. Desrosiers, F. Soumis. 1991. The pickup and delivery problem with time windows. Eur. J. Oper. Res. 54 7–22.

	\bibitem{root}
	Gendreau, M., F. Guertin, J.-Y. Potvin, R. Séguin. 1998. Neighbor- hood search heuristics for a dynamic vehicle dispatching prob- lem with pick-ups and deliveries. Tech. Rep. CRT-98-10, Centre de Recherche sur les Transport, Universite de Montreal, Mon- treal, Canada.

	\bibitem{root}
	Lau, H. C., Z. Liang. 2001. Pickup and delivery with time win- dows: Algorithms and test case generation. ICTAI-2001, 13th IEEE Conf. Tools Artificial Intelligence, Dallas, TX, 333–340.

	\bibitem{root}
	Li, H., A. Lim. 2001. A metaheuristic for the pickup and delivery problem with time windows. ICTAI-2001, 13th IEEE Conf. Tools Artificial Intelligence, Dallas, TX, 160–170.

	\bibitem{root}
	Lim, H., A. Lim, B. Rodrigues. 2002. Solving the pick up and deliv- ery problem with time windows using “squeaky wheel” opti- mization with local search. Technical Report 2002, Paper 7- 2002, Singapore Management University, Singapore.

	\bibitem{root}
	Martello, S., P. Toth. 1981. An algorithm for the generalized assign- ment problem. J. P. Brans, ed. Operational Research '81. North- Holland, New York.
	Mladenovic', N., P. Hansen. 1997. Variable neighborhood search. Comput. Oper. Res. 24(11) 1097–1100.

	\bibitem{root}
	Nanry, W. P., J. W. Barnes. 2000. Solving the pickup and delivery problem with time windows using reactive tabu search. Trans- portation Res. Part B 34 107–121.

	\bibitem{root}
	Pisinger, D., S. Ropke. 2005. A general heuristic for vehicle routing problems. Comput. Oper. Res. Forthcoming. http:// www.sciencedirect.com
	Potvin, J.-Y., J.-M. Rousseau. 1993. A parallel route building algo- rithm for the vehicle routing and scheduling problem with time windows. Eur. J. Oper. Res. 66 331–340.

	\bibitem{root}
	Ropke, S. 2002. Heuristics for the multi-vehicle pickup and deliv- ery problem with time windows. Masters Thesis 01-11-8, Department of Computer Science, University of Copenhagen, Denmark.

	\bibitem{root}
	Ropke, S., D. Pisinger. 2004. An adaptive large neighborhood search heuristic for the pickup and delivery problem with time win- dows. Technical report, Department of Computer Science, Uni- versity of Copenhagen, Denmark.

	\bibitem{root}
	Ropke, S., D. Pisinger. 2006. A unified heuristic for a large class of vehicle routing problems with backhauls. Eur. J. Oper. Res. 171 750–775.
	Shaw, P. 1997. A new local search algorithm providing high quality solutions to vehicle routing problems. Technical report, Depart- ment of Computer Science, University of Strathclyde, Scotland.

	\bibitem{root}
	Shaw, P. 1998. Using constraint programming and local search methods to solve vehicle routing problems. Proc. CP-98 (Fourth Internat. Conf. Principles Practice Constraint Programming).

	\bibitem{root}
	Schrimpf, G., J. Schneider, H. Stamm-Wilbrandt, G. Dueck. 2000. Record breaking optimization results using the ruin and recre- ate principle. J. Comput. Phys. 159 139–171.

	\bibitem{root}
	Sigurd, M., D. Pisinger, M. Sig. 2004. The pickup and delivery prob- lem with time windows and precedences. Transportation Sci. 38 197–209.

	\bibitem{root}
	SINTEF. Vehicle routing and travelling salesperson problems. http://www.sintef.no/static/am/opti/projects/top/vrp/ benchmarks.html.

	\bibitem{root}
	Solomon, M. M. 1987. Algorithms for the vehicle routing and scheduling problems with time window constraints. Oper. Res. 35 254–265.

	\bibitem{root}
	Toth, P., D. Vigo. 2002. An overview of vehicle routing problems. P. Toth, D. Vigo, eds. The Vehicle Routing Problem. SIAM Mono- graphs on Discrete Mathematics and Applications, Vol. 9. SIAM, Philadelphia, PA, 1–26.

	\bibitem{root}
	Trick, M. A. 1992. A linear relaxation heuristic for the generalized assignment problem. Naval Res. Logist. 39 137–152.

	\bibitem{root}
	Xu, H., Z.-L. Chen, S. Rajagopal, S. Arunapuram. 2003. Solving a practical pickup and delivery problem. Transportation Sci. 37 347–364.
\end{thebibliography}